---
title: "Linking Health Indicators to Refinery Emissions"
author: "Niklas Lollo"
output: 
  html_document:
    toc: true
runtime: shiny
---

```{r setup, include=FALSE, echo=F}
# http://rmarkdown.rstudio.com/index.html
knitr::opts_chunk$set(echo = TRUE)
# List of R packages
required_pkg <- c("tidyverse","maptools", "rgdal", "spatstat", "ggmap")
#pkgs_not_installed <- required_pkg[!sapply(required_pkg, function(p) require(p, character.only=T))]
#install.packages(pkgs_not_installed, dependencies=TRUE)

# Load all libraries at once.
lapply(required_pkg, library, character.only = TRUE) 
```

```{r time dataset, message=F, warning=F, echo=F}
# Inputs here are date_begin and date_end
timeframe <- function(date_begin = "2016-05-09 14:00", date_end = "2016-08-11 0:00"){
  exp_begin = as.POSIXct(date_begin, tz="GMT", format="%Y-%m-%d %H")
  exp_end = as.POSIXct(date_end, tz="GMT", format="%Y-%m-%d %H")
  out = data_frame(day = seq(exp_begin, exp_end, by = 3600))
  return(out)
}
full_time <- timeframe()
# Manual
date_begin = "2016-05-09 14:00"
date_end = "2016-08-11 0:00"
exp_begin = as.POSIXct(date_begin, tz="GMT", format="%Y-%m-%d %H")
exp_end = as.POSIXct(date_end, tz="GMT", format="%Y-%m-%d %H")
```

```{r load air quality data, warning=F, message=F, echo=F}
# load air quality data
feed_4902 <- read_csv("../refinery_data/airQuality_feed4902_data.csv")
feed_4901 <- read_csv("../refinery_data/airQuality_feed4901_data.csv")
feed_4902_methane <- read_csv("../refinery_data/airQuality_N20_methane_4902_data.csv")

# Make hourly averages
## Set zero's to NA to not "mess up" average
feed_4902[feed_4902 == 0] <- NA
## Create new dataframe
hourly_4902 <- feed_4902 %>%
  # Select the day and hour of interest (get rid of minute and seconds)
  group_by(
    # Creates new variables
    day = as.POSIXct(cut(t, breaks='hour'))) %>%
  summarise(
    # Averages
    sulfur_dioxide = mean(sulfurDioxide, na.rm=T),
    carbon_monoxide = mean(carbonMonoxide, na.rm=T),
    ozone = mean(ozone, na.rm=T)) %>%
  ungroup() %>%
  # Arrange the columns by day
  arrange(day) %>%
  full_join(full_time, by = "day")

# Do the same for the other feed4902 dataframe
feed_4902_methane[feed_4902_methane == 0] <- NA
hourly_4902_methane <- feed_4902_methane %>%
    group_by(
    # Creates new variables
    day = as.POSIXct(cut(t, breaks='hour'))) %>%
  summarise(
    # Averages
    methane = mean(Methane, na.rm=T),
    benzene = mean(Benzene, na.rm=T),
    nitrous_oxide = mean(`Nitrous Oxide`, na.rm=T)) %>%
  # Arrange the columns by day
  arrange(day) %>%
  full_join(full_time, by = "day")

air_qual_4902 <- right_join(hourly_4902, hourly_4902_methane, by = "day") %>%
  mutate(id = "feed_4902")

# Do the same for feed 4901
feed_4901[feed_4901 == 0] <- NA
air_qual_4901 <- feed_4901 %>%
  # Select the day and hour of interest (get rid of minute and seconds)
  group_by(
    # Creates new variables
    day = as.POSIXct(cut(t, breaks='hour'))) %>%
    # Averages
  summarise(
    sulfur_dioxide = mean(sulfurDioxide, na.rm=T),
    carbon_monoxide = mean(carbonMonoxide, na.rm=T),
    ozone = mean(ozone, na.rm=T)) %>%
  mutate(
    id = "feed_4901"
  ) %>%
  # Arrange the columns by day
  arrange(day) %>%
  full_join(full_time, by = "day")

# Remove intermediate dataframes
rm(hourly_4902, hourly_4902_methane)
# Remove original dataframes
rm(feed_4901, feed_4902,feed_4902_methane)
```

```{r load paco data, warning=F, message=F, echo=F}
# load paco data
paco <- read_csv("../refinery_data/paco_all.csv")

# Creating new dataframe
paco <- paco %>%
  mutate(
    # Remove extra values (supposed to be a correction)
    when=gsub("\\+0000","",when),
    # Converting / to - part 1
    when=sub("[[:punct:]]","-",when),
    # Converting / to - part 2
    when=sub("\\/","-",when),
    # Rename persons of interest
    who = replace(who, who=="meaningfrommonitoring3@gmail.com", "m3"),
    who = replace(who, who=="meaningfrommonitoring5@gmail.com", "m5"),
    who = replace(who, who=="meaningfrommonitoring7@gmail.com", "m7"),
    who = replace(who, who=="meaningfrommonitoring9@gmail.com", "m9"),
    who = replace(who, who=="meaningfrommonitoring10@gmail.com", "m10"),
    who = replace(who, who=="meaningfrommonitoring11@gmail.com", "m11"),
    who = replace(who, who=="meaningfrommonitoring16@gmail.com", "m16")
  ) %>%
  mutate(
    # Make person column as a factor
    id = as.factor(who),
    # Convert date column to datetime
    t = as.POSIXct(when, tz="GMT", format="%Y-%m-%d %H:%M:%S")) %>%
  # Remove columns not of interest
  select(-c(who, when, appId, pacoVersion, experimentId,
            experimentName, experimentVersion, experimentGroupName,
            actionTriggerId, actionId, actionSpecId, responseTime,
            scheduledTime, timeZone)) %>%
  # Select persons of interest 
  filter(id == "m3" | id == "m5" | id == "m7" |
      id == "m9" | id == "m10" | id == "m11" |
      id == "m16") %>%
  arrange(t)

# Remove incorrect values
paco$SPO2[paco$SPO2 == 0 | paco$SPO2 > 100 | paco$SPO2 < 80] <- NA
# Make hourly dataframe
paco_hourly <- paco %>%
  # Select the person, day and hour of interest
  group_by(id,
    # Creates new variables
    day = as.POSIXct(cut(t, breaks='hour'))) %>%
  # Average
  summarise(blood_oxygen = mean(SPO2, na.rm=T)) %>%
  filter(is.na(blood_oxygen) == F) %>% 
  complete(expand(nesting(id), day = seq(exp_begin, exp_end, by = 3600)),  
           #completing all levels of id:day
              fill = list(blood_oxygen = NA)) %>%
  # Select persons of interest 
  filter(id == "m3" | id == "m5" | id == "m7" |
      id == "m9" | id == "m10" | id == "m11" |
      id == "m16") %>%
  # Arrange the columns by person
  arrange(id)

rm(paco)
```

```{r load fitbit data, message=F, warning=F, echo=F}
# Load fitbit data
fb_body <- read_csv("../refinery_data/fitbit_body_data.csv")
fb_activities <- read_csv("../refinery_data/fitbit_activities_data.csv")
fb_sleep <- read_csv("../refinery_data/fitbit_sleep_data.csv")

# Merge daily fitbit data
fb_daily <- inner_join(fb_body, fb_activities, fb_sleep, by = "t")
rm(fb_body, fb_activities, fb_sleep)

# Load intraday activity
fb_intraday_m3 <- read_csv("../refinery_data/fitbit_intradayactivities_m3.csv") %>%
  mutate(id = "m3")
fb_intraday_m5 <- read_csv("../refinery_data/fitbit_intradayactivities_m5.csv")%>%
  mutate(id = "m5")
fb_intraday_m7 <- read_csv("../refinery_data/fitbit_intradayactivities_m7.csv")%>%
  mutate(id = "m7")
fb_intraday_m9 <- read_csv("../refinery_data/fitbit_intradayactivities_m9.csv")%>%
  mutate(id = "m9")
fb_intraday_m11 <- read_csv("../refinery_data/fitbit_intradayactivities_m11.csv")%>%
  mutate(id = "m11")
fb_intraday_m16 <- read_csv("../refinery_data/fitbit_intradayactivities_m16.csv")%>%
  mutate(id = "m16")

# Combine dataframes
fb_intra <- bind_rows(fb_intraday_m3, fb_intraday_m5, .id = NULL) %>%
  bind_rows(fb_intraday_m7, .id = NULL) %>%
  bind_rows(fb_intraday_m9, .id = NULL) %>%
  bind_rows(fb_intraday_m11, .id = NULL) %>%
  bind_rows(fb_intraday_m16, .id = NULL) %>%
  mutate(id = as.factor(id))

# Remove intermediate columns
rm(fb_intraday_m3, fb_intraday_m5, fb_intraday_m7,
   fb_intraday_m9, fb_intraday_m11, fb_intraday_m16)

# Make hourly data
fb_intraday <- fb_intra %>% 
  group_by(id,
    # Creates new variables
    day = as.POSIXct(cut(t, breaks='hour'))) %>%
  summarise(
    # Averages
    calories = sum(calories, na.rm=T),
    distance = sum(distance, na.rm=T),
    steps = sum(steps, na.rm=T),
    heart_rate = mean(heart, na.rm=T),
    floors = sum(floors, na.rm=T),
    elevation = sum(elevation, na.rm=T)) %>%
    complete(expand(nesting(id), day = seq(exp_begin, exp_end, by = 3600)),  
         #completing all levels of id:day
              fill = list(calories = NA,distance = NA,
                          steps = NA, heart_rate = NA,
                          floors = NA, elevation = NA)) %>%
  # Select persons of interest 
  filter(id == "m3" | id == "m5" | id == "m7" |
      id == "m9" | id == "m10" | id == "m11" |
      id == "m16") %>%
  # Arrange the columns by day
  arrange(id, day)

# Remove original dataframes
rm(fb_intra, fb_daily)
```

```{r make individual data, message=F, warning=F, echo=F}
# Make demographics dataframe
demographics <- data_frame(
  id =  c("m3","m5","m7","m9","m10","m11","m16"),
  age = c(21,34,45,33,56,67,51),
  sex = as.factor(c("male","female","male","female","female","male","female")))

# Merge with fb_intraday
fb_data <- left_join(fb_intraday, demographics, by = "id") %>%
  ungroup %>%
  mutate(
    id = as.factor(id)
  )
paco_data <- full_join(paco_hourly, demographics, by = "id") %>%
  ungroup %>%
  mutate(
    id = as.factor(id)
    )

# Individual data
individual_data <- fb_data %>%
  bind_rows(paco_data, .id = NULL)

# Remove unnecessary dataframes
rm(paco_hourly, fb_intraday)
```

```{r create air quality dataset, message=F, warning=F, echo=F}
# Air Quality data
air_quality <- full_join(full_time, air_qual_4901, by = "day") %>%
  full_join(air_qual_4902, 
            by = c("day", "id", "sulfur_dioxide",
                   "carbon_monoxide", "ozone")) %>%
  select(day, id, everything()) %>%
  mutate(id = as.factor(id)) %>%
  arrange(day)
```

```{r moving window, message=F, warning=F, echo=F}
# Set hours of interest
hours = 8
# Moving window average of sulfur dioxide
air_quality <- 
  air_quality %>% 
  group_by(id) %>% 
  arrange(day) %>%
  mutate(
    sulfur_dioxide_exposure_window = 
      zoo::rollapply(sulfur_dioxide,
                     width = hours, #how many data to include
                     FUN = mean, #function is mean
                     na.rm = T, #avoids NA
                     partial = T, #skips unnecessary datapoints
                     align = "right")) %>%
  ungroup %>%
  arrange(id, day)
```

```{r TO DO filter your data, message=F, warning=F, echo=F}
ind_data_filtered <- individual_data %>%
  filter(heart_rate > 30 & heart_rate < 200 | is.na(heart_rate)) %>% # put bounds on heart rate
  filter(sex == "male" & age > 50) # Select demographic filters

# Make the final dataframe (do not change)
final_df <- ind_data_filtered %>%
  bind_rows(air_quality, .id = NULL) %>%
  mutate(id = as.factor(id))

# Make sure to write your note!
filter_statement <- paste0("NOTE: This analysis looks at men over 50.")
cat(filter_statement)
```
# Notes on Data Sources
Methane and sulfur dioxide readings collected in real-time by [Fenceline.org](fenceline.org)

Heart rate collected with Fitbit health monitors

Blood Oxygen levels recorded with a pulse oximeter^[[Pulse Oximetry wiki](https://en.wikipedia.org/wiki/Pulse_oximetry)]

# Correlations

## Plots^[For help on interpreting a correlation plot, visit this [page](https://www.mathsisfun.com/data/images/correlation-examples.svg)]

t-tests for Statistical Significance^[Welch two-sample t-test]

These tests compare the difference between the population mean health outcome (e.g. heart rate) and the mean health outcome when the environmental input (e.g. methane) is above its median value.
```{r plots, warning=F, message=F, echo=F}
# http://zevross.com/blog/2014/08/04/beautiful-plotting-in-r-a-ggplot2-cheatsheet-3/#add-text-annotation-in-the-top-right-top-left-etc.-annotation_custom-and-friends

# Create ggplot theme
theme_fair_tech <- theme(
  legend.position = "bottom",
  panel.background = element_rect(fill = NA),
  # panel.border = element_rect(fill = NA, color = "grey75"),
  axis.ticks = element_line(color = "grey95", size = 0.3),
  panel.grid.major = element_line(color = "grey95", size = 0.3),
  panel.grid.minor = element_line(color = "grey95", size = 0.3),
  legend.key = element_blank(),
  plot.title = element_text(size=20, face="bold", color = "maroon", 
    margin = margin(10, 0, 10, 0))
  )

my_plot <- function(df = final_df,
                    time_var = day,
                    health_var,
                    health_var_name,
                    health_var_units,
                    pollutant_var,
                    pollutant_var_name, 
                    pollutant_var_units){
  
  health_var <- enquo(health_var)
  pollutant_var <- enquo(pollutant_var)
  time_var <- enquo(time_var)

  # Summary dataframe
  df %>%
    select(!!health_var, !!pollutant_var, !!time_var) %>%
    filter(!is.na(!!health_var) | !is.na(!!pollutant_var))%>%
    group_by(!!time_var) %>%
    summarize(mean_health_var = mean(!!health_var, na.rm=T),
              mean_pollutant_var = mean(!!pollutant_var, na.rm=T)) %>% 
    ungroup %>%
    filter(!is.na(mean_health_var) & !is.na(mean_pollutant_var)) ->
    df_new
  
  # Correlation
  df_new %>%
    select(mean_health_var, mean_pollutant_var) %>% 
    boot::corr() %>%
    round(3) -> correlation
  
  # t-test
  df %>% 
    transmute(health = !!health_var) -> health
  
  df_new %>%
    filter(mean_pollutant_var > median(mean_pollutant_var)) %>%
    select(mean_health_var) -> sample
  # Perhaps should compare to air_quality standards instead of median
  t_test <- t.test(x = sample[[1]], y = health[[1]])

  # Plot
  df_new %>%
    ggplot(aes(x= mean_health_var, y= mean_pollutant_var)) + 
      geom_point() +
      geom_smooth(method = "lm", se=F) +
      theme_fair_tech + 
      labs(x=paste0(health_var_name," (",health_var_units,")"), 
           y=paste0(pollutant_var_name," (",pollutant_var_units,")")) +
      ggtitle(paste0(health_var_name, " & ", pollutant_var_name),
              subtitle=paste0("r = ", correlation,
                              ", p-value = ",round(t_test$p.value,4)))
}

# Blood Oxygen and Sulfur Dioxide
my_plot(health_var = blood_oxygen,
        health_var_name = "Blood Oxygen",
        health_var_units = "%",
        pollutant_var = sulfur_dioxide_exposure_window,
        pollutant_var_name = "Sulfur Dioxide",
        pollutant_var_units = "ppb")

# Heart Rate and Sulfur Dioxide
my_plot(health_var = heart_rate,
        health_var_name = "Heart Rate",
        health_var_units = "bpm",
        pollutant_var = sulfur_dioxide_exposure_window,
        pollutant_var_name = "Sulfur Dioxide",
        pollutant_var_units = "ppb")

# Heart Rate and Methane
my_plot(health_var = heart_rate,
        health_var_name = "Heart Rate",
        health_var_units = "bpm",
        pollutant_var = methane,
        pollutant_var_name = "Methane",
        pollutant_var_units = "ppb")
```

```{r shiny}
#final_df %>%
#  group_by(day) %>%
#  summarize(
#    methane = mean(methane, na.rm=T),
#    heart_rate = mean(heart_rate, na.rm=T)
#  ) %>%
#  ungroup %>%
#  filter(!is.na(methane) & !is.na(heart_rate)) %>%
#  ggplot() + 
#    geom_point(aes(x = methane, y = heart_rate))

library(rsconnect)
server <- function(input, output) {
  
  output$healthSelect <- renderUI({
    selectInput("health_var", "Health Indicator", 
                choices = c("heart_rate", "blood_oxygen"), 
                selected = "heart_rate")
  })
  
  output$pollutantSelect <- renderUI({
    selectInput("pollutant_var", "Pollutant", 
                choices = c("sulfur_dioxide_exposure_window", "methane"),
                selected = "methane")
  })
  
  output$plot1 <- renderPlot({
    final_df %>%
      group_by(day) %>%
      summarize(
        "pollutant" = mean(UQ(input$pollutant_var), na.rm=T),
        "health" = mean(UQ(input$health_var), na.rm=T)
      ) %>%
      ungroup -> new_df
    ggplot(new_df) +
      geom_point(aes_string(x = "pollutant", y = "health"))
  })
  
}

ui <- fluidPage(
  sidebarLayout(
    sidebarPanel(
      h3('Indicators'),
      uiOutput('healthSelect'),
      uiOutput("pollutantSelect")
    ),
    mainPanel(
      plotOutput('plot1'),
      uiOutput('table')
    )
  )
)

shinyApp(ui = ui, server = server)
```
