---
title: "Linking Health Indicators to Refinery Emissions"
author: "Niklas Lollo"
output: 
  html_document:
    toc: true
---

```{r setup, include=FALSE, echo=F}
knitr::opts_chunk$set(echo = TRUE)
# List of R packages
required_pkg <- c("tidyverse","maptools", "rgdal", "spatstat", "ggmap", "geosphere", "leaflet", "spdep", "sp")
#pkgs_not_installed <- required_pkg[!sapply(required_pkg, function(p) require(p, character.only=T))]
#install.packages(pkgs_not_installed, dependencies=TRUE)

# Load all libraries at once.
lapply(required_pkg, library, character.only = TRUE) 
```

```{r load data, warning=F, message=F, echo=F}
individual_data <- read_csv("data/individual_data.csv")
air_quality <- read_csv("data/air_quality.csv")
```

```{r TO DO filter your data, message=F, warning=F, echo=F}
# Merged dataframe 
total_df <- individual_data %>%
  #filter(sex == "male" & age > 50) %>% # TO DO: select demographic filters
  bind_rows(air_quality, .id = NULL) %>%
  mutate(id = as.factor(id),
         methane = as.numeric(methane),
         sex = as.factor(sex),
         day = as.POSIXct(day),
         blood_oxygen = as.numeric(blood_oxygen),
         sym_factor = as.factor(symptoms),
         doctor = as.factor(doctor),
         medicine = as.factor(medicine),
         symptoms = ifelse(sym_factor %in% c(1:9), 1,
                           ifelse(is.na(sym_factor), NA, 0)),
         benzene = as.numeric(benzene),
         nitrous_oxide = as.numeric(nitrous_oxide),
         xylene = as.numeric(xylene)
         )

# Intermediate dataframe - remove pollution variables
binding_df <- total_df %>%
  select(-c(methane, sulfur_dioxide, ozone,
            benzene, pm_2.5, carbon_monoxide,
            sulfur_dioxide_exposure_window, xylene,
            nitrous_oxide))

# To be used in spatial analysis
spatial_df <- total_df %>%
  group_by(day) %>%
  summarize(
    methane = mean(methane, na.rm=T),
    sulfur_dioxide = mean(sulfur_dioxide_exposure_window, na.rm=T),
    ozone = mean(ozone, na.rm=T),
    benzene = mean(benzene, na.rm=T),
    pm_2.5 = mean(pm_2.5, na.rm=T),
    carbon_monoxide = mean(carbon_monoxide, na.rm=T),
    so2_not_window = mean(sulfur_dioxide, na.rm=T),
    xylene = mean(xylene, na.rm=T),
    nitrous_oxide = mean(nitrous_oxide, na.rm=T)
  ) %>%
  ungroup %>%
  right_join(binding_df, by = "day") %>%
  filter(!is.na(id)) %>%
  arrange(day)

# Remove intermediate dataframe
rm(binding_df)

# Create data_frame for standard analysis
summary_df <- total_df %>%
  group_by(day) %>%
  summarize(
    # Pollutants
    methane = mean(methane, na.rm=T),
    sulfur_dioxide = mean(sulfur_dioxide_exposure_window, na.rm=T),
    ozone = mean(ozone, na.rm=T),
    benzene = mean(benzene, na.rm=T),
    pm_2.5 = mean(pm_2.5, na.rm=T),
    carbon_monoxide = mean(carbon_monoxide, na.rm=T),
    so2_not_window = mean(sulfur_dioxide, na.rm=T),
    xylene = mean(xylene, na.rm=T),
    nitrous_oxide = mean(nitrous_oxide, na.rm=T),
    # Health indicators
    heart_rate = mean(heart_rate, na.rm=T),
    blood_oxygen = mean(blood_oxygen, na.rm=T),
    # Summary demographics
    time_in_area = mean(time_in_area, na.rm=T),
    age = mean(age, na.rm=T),
    home_dist_to_refinery = mean(home_dist_to_refinery, na.rm=T),
    symptoms = mean(symptoms, na.rm=T),
    text_entry = first(text_entry)
    ) %>%
  ungroup
```

## Filters
This analysis looks at 

```{r plot theme, warning=F, message=F, echo=F}
# Create plotting theme
theme_fair_tech <- theme(
  legend.position = "bottom",
  panel.background = element_rect(fill = NA),
  # panel.border = element_rect(fill = NA, color = "grey75"),
  axis.ticks = element_line(color = "grey95", size = 0.3),
  panel.grid.major = element_line(color = "grey95", size = 0.3),
  panel.grid.minor = element_line(color = "grey95", size = 0.3),
  legend.key = element_blank(),
  plot.title = element_text(size=20, face="bold", color = "maroon", 
    margin = margin(10, 0, 10, 0))
  )
```

# Correlations

## Plots^[For help on interpreting a correlation plot, visit this [page](https://www.mathsisfun.com/data/images/correlation-examples.svg)]

t-tests for Statistical Significance^[Welch two-sample t-test]

These tests compare the difference between the population mean health outcome (e.g. heart rate) and the mean health outcome when the environmental input (e.g. methane) is above its median value. Actually does not do this anymore.

+ To do: should make this analysis (maybe also) look at when pollutant is above permitted limit or if it is present at all.

```{r make plot function, warning=F, message=F, echo=F}
my_plot <- function(df = summary_df,
                    health_var,
                    health_var_name,
                    health_var_units,
                    pollutant_var,
                    pollutant_var_name, 
                    pollutant_var_units){
  
  health_var <- enquo(health_var)
  pollutant_var <- enquo(pollutant_var)

  # Correlation
  df %>%
    filter(!is.na(!!health_var) & !is.na(!!pollutant_var)) %>%
    select(UQ(health_var), UQ(pollutant_var)) %>% 
    boot::corr() %>% # why not use cor()?
    round(3) -> correlation

  # t-test
  df %>% 
    filter(!is.na(!!health_var)) %>%
    transmute(health = !!health_var) -> health
  
  df %>%
    filter(!is.na(!!health_var) & !is.na(!!pollutant_var)) %>%
    # Not filtering for above median anymore
    select(UQ(health_var)) -> sample

  # Perhaps should compare to air_quality standards instead of median
  t_test <- t.test(x = sample[[1]], y = health[[1]])

  df %>%
    filter(!is.na(!!health_var) & !is.na(!!pollutant_var)) %>%
    transmute(
      health = !!health_var,
      pollutant = !!pollutant_var) -> df_new

    # Plot
  df_new %>%
    ggplot() + 
      geom_point(aes(x= health, y= pollutant)) +
      geom_smooth(aes(x= health, y= pollutant), 
                  method = "lm", se=F) +
      scale_x_continuous() +
      scale_y_continuous() +
      theme_fair_tech + 
      labs(x=paste0(health_var_name,
                    " (",health_var_units,")"), 
           y=paste0(pollutant_var_name,
                    " (",pollutant_var_units,")")) +
      ggtitle(paste0(health_var_name, 
                     " & ", pollutant_var_name),
              subtitle=paste0("r = ", correlation,
                              ", p-value = ",
                              round(t_test$p.value,4)))
}
```

```{r run plots, warning=F, message=F, echo=F}
# Blood Oxygen and Sulfur Dioxide
my_plot(health_var = blood_oxygen,
        health_var_name = "Blood Oxygen",
        health_var_units = "%",
        pollutant_var = sulfur_dioxide,
        pollutant_var_name = "Sulfur Dioxide",
        pollutant_var_units = "ppb")

# Heart Rate and Sulfur Dioxide
my_plot(health_var = heart_rate,
        health_var_name = "Heart Rate",
        health_var_units = "bpm",
        pollutant_var = sulfur_dioxide,
        pollutant_var_name = "Sulfur Dioxide",
        pollutant_var_units = "ppb")

# Heart Rate and Methane
my_plot(health_var = heart_rate,
        health_var_name = "Heart Rate",
        health_var_units = "bpm",
        pollutant_var = methane,
        pollutant_var_name = "Methane",
        pollutant_var_units = "ppb")
```

# Spatial Analysis

## Plot points
```{r plot spatial points, echo=F, message=F, warning=F}
# Load contra costa county shapefile
cc <- readOGR(dsn ="data/location/BaseMapGrid", 
              layer = "GRD_PWD_BaseMap_0106",
              verbose = F)

# Convert to longlat data
cc_wgs84 <- spTransform(cc, 
                        CRS("+proj=longlat +datum=WGS84"))


# Basic plot of contra costa county
r <- ggplot(cc_wgs84, aes(x=long, y=lat)) + 
  geom_path(aes(group=group)) + 
  coord_map("mercator") + 
  geom_point(aes(x= air_quality$longitude[[1]], y= air_quality$latitude[[1]])) 

# Plot ids on contra costa county map
r + 
  geom_point(data=spatial_df, 
             aes(x=longitude, y=latitude, 
                 color = id), 
             na.rm = T)

# Plot heart rate over space
r + 
  geom_point(data=spatial_df, 
             aes(x=longitude, y=latitude,
                 color = heart_rate), 
             na.rm = T)
```

## Autocorrelation
```{r spatial autocorrelation, echo=F, warning=F, message=F}
### Dtermine which points are near
# Select only the spatial points
points <- spatial_df %>% select(longitude, latitude)

# Turn into spatial dataframe
spat_points <- SpatialPoints(points,
                             proj4string = CRS("+proj=longlat +datum=WGS84"))

# Nearest neighbor assignment
points_nb <- dnearneigh(spat_points, d1 = 0, d2 = .1)

# Convert into weights
spatial_weights <- nb2mat(points_nb, zero.policy = T)

# Convert into listw
spatial_listw <- nb2listw(points_nb, zero.policy =T)

# Intermediate dataframe removing LatLong
spatial_df_t <- spatial_df %>% select(-c(longitude, latitude))

# Make spatial points dataframe with merge
spatial_points <- SpatialPointsDataFrame(coords = points,
                            data = spatial_df_t,
                            proj4string = 
                              CRS("+proj=longlat +datum=WGS84"),
                            match.ID = T)

# Moran's I Monte Carlo
moran_out <- moran.mc(spatial_df$heart_rate,
                      spatial_listw,
                      nsim=99,
                      zero.policy = T,
                      na.action = na.omit)

paste0("Moran's I p-value: ", moran_out$p.value)
paste0("Moran's I z-score: ", moran_out$statistic)
```

## Load Cal Enviro Screen
```{r Cal Enviro Screen, warning=F, echo=F, message=F}
## Load community health assessment data
ces <- readOGR(dsn ="data/location/CES3Results", 
               layer = "CES3Results", 
               verbose = F)

# Spatial transform
ces_lat <- spTransform(ces, CRS("+proj=longlat +datum=WGS84"))

# Subset for Contra Costa County
ces_lat_cc <- subset(ces_lat, County == "Contra Costa")
# head(ces_lat_cc@data) # Examine the data of interest

# Add a new column "id" from the rownames
ces_lat_cc@data$id <- rownames(ces_lat_cc@data)

# Create a dataframe from the spatial object
new_data <- fortify(ces_lat_cc, region = "id")

# Merge the "fortified" data with the spatial object data
waterDF <- merge(new_data, ces_lat_cc@data, by = "id")

# Map (e.g. the cardiovascular scores)
ggplot(waterDF, 
       aes(x=long, y=lat,
           group = group,
           fill = Cardiovasc)) +
  geom_polygon()  +
  geom_path(color = "white") +
  scale_fill_gradient(
    low = "plum1", high = "purple4",
    breaks=c(500000,1000000,1500000),
    labels=c("Low","Medium","High")) +
  coord_map("mercator") + 
  ggtitle("Cardiovascular Health")
```
